When two-point correlation function fails to describe a structure...

### 2024-03-07 paper.tex
S₀ → Id

By looking at \cref{eq:s2z} it becomes clear that the following two
substitutions, designated $S_k$ and $R$, do not change $s_2(z)$ and hence, the
two-point correlation function:
\begin{equation}
  \begin{aligned}
    R&: f(z) \rightarrow f(z^{-1}) \\
    S_k&: f(z) \rightarrow z^{-k}f(z)
  \end{aligned}
  \label{eq:group-elts}
\end{equation}
The set
\begin{equation}
  G = \{R, S_0, S_1, S_{-1}, \dots \}
  \label{eq:group}
\end{equation}
and a binary operation $\circ$ (composition) form a group of transforms of $a$
which do not change the two-point function. Its neutral element is $S_0$ and the
associativity and invertibility are trivial to prove.

@@

A reciprocal polynomial of $f(x)$ has the same degree as $f$. Suppose $f$
factors into $M$ different factors in $\mathbb{Z}[x]$:
\begin{equation*}
  f(x) = \prod_{i=1}^M f_i(x)
\end{equation*}
The image of the two-point function is then
\begin{equation*}
  s(x) = f(x)f(x^{-1}) = \prod_{i=1}^M f_i(x) \prod_{i=1}^M f_i(x^{-1})
\end{equation*}
Let now select all non-palindrome factors $f_{r_1}, f_{r_2}, \dots, f_{r_L}$ of
$f$, $0 \le L \le M$ and replace them with their reciprocals. Obviously, the
resulting correlation function won't change:
\begin{equation*}
  s(x) = \underbrace{\prod_{i=1}^L x^{\deg f_{r_i}} f_{r_i}(x^{-1}) \prod_{i=1}^{M-L} f_{o_i}(x)}_{\tilde{f}(x)}
  \underbrace{\prod_{i=1}^L x^{-\deg f_{r_i}} f_{r_i}(x) \prod_{i=1}^{M-L} f_{o_i}(x^{-1})}_{\tilde{f}(x^{-1})}
\end{equation*}
where $o_1, o_2, \dots, o_{M-L}$ run through indices of original (not replaced)
factors. Now recombining factors into $\tilde{f}(x)$ and $\tilde{f}(x^{-1})$ we
obtain a polynomial $\tilde{f}$ with the same degree as $f$ and $f(0) = \tilde{f}(0)$.
Now note that with this additional demand on the free coefficient, we can
restrict the group of transforms \cref{eq:group} to a subgroup $\{S_0, R\}$
because a non-zero shift will change the degree of $f$ which is not allowed. We
have two possible outcomes regarding $\tilde{f}$:
\begin{itemize}
\item Only at most one factor of $\tilde{f}$ is not a palindrome
  ($L \le 1$). This gives us $\tilde{f}(x) = f(x^{-1})$ which is a trivial
  collision.
\item At least two factors of $\tilde{f}$ are not palindromes ($L > 1$). In
  this case $\tilde{f} \ne f$ and cannot be obtained from $f$ using transforms
  in the subgroup $\{S_0, R\}$.
\end{itemize}

>>>>

By looking at \cref{eq:s2z} it becomes clear that the following two
substitutions, designated $S_k$ and $R$, do not change $s_2(z)$ and hence, the
two-point correlation function:
\begin{equation}
  \begin{aligned}
    R&: f(z) \rightarrow f(z^{-1}) \\
    S_k&: f(z) \rightarrow z^{-k}f(z)
  \end{aligned}
  \label{eq:group-elts}
\end{equation}
The set
\begin{equation}
  G = \{R, Id, S_1, S_{-1}, \dots \}
  \label{eq:group}
\end{equation}
and a binary operation $\circ$ (composition) form a group of transforms of $a$
which do not change the two-point function. Its neutral element $Id = S_0$ and
the associativity and invertibility are trivial to prove.

@@

A reciprocal polynomial of $f(x)$ has the same degree as $f$. Suppose $f$
factors into $M$ different factors in $\mathbb{Z}[x]$:
\begin{equation*}
  f(x) = \prod_{i=1}^M f_i(x)
\end{equation*}
The image of the two-point function is then
\begin{equation*}
  s(x) = f(x)f(x^{-1}) = \prod_{i=1}^M f_i(x) \prod_{i=1}^M f_i(x^{-1})
\end{equation*}
Let now select all non-palindrome factors $f_{r_1}, f_{r_2}, \dots, f_{r_L}$ of
$f$, $0 \le L \le M$ and replace them with their reciprocals. Obviously, the
resulting correlation function won't change:
\begin{equation*}
  s(x) = \underbrace{\prod_{i=1}^L x^{\deg f_{r_i}} f_{r_i}(x^{-1}) \prod_{i=1}^{M-L} f_{o_i}(x)}_{\tilde{f}(x)}
  \underbrace{\prod_{i=1}^L x^{-\deg f_{r_i}} f_{r_i}(x) \prod_{i=1}^{M-L} f_{o_i}(x^{-1})}_{\tilde{f}(x^{-1})}
\end{equation*}
where $o_1, o_2, \dots, o_{M-L}$ run through indices of original (not replaced)
factors. Now recombining factors into $\tilde{f}(x)$ and $\tilde{f}(x^{-1})$ we
obtain a polynomial $\tilde{f}$ with the same degree as $f$ and $f(0) = \tilde{f}(0)$.
Now note that with this additional demand on the free coefficient, we can
restrict the group of transforms \cref{eq:group} to a subgroup $\{Id, R\}$
because a non-zero shift will change the degree of $f$ which is not allowed. We
have two possible outcomes regarding $\tilde{f}$:
\begin{itemize}
\item Only at most one factor of $\tilde{f}$ is not a palindrome
  ($L \le 1$). This gives us $\tilde{f}(x) = f(x^{-1})$ which is a trivial
  collision.
\item At least two factors of $\tilde{f}$ are not palindromes ($L > 1$). In
  this case $\tilde{f} \ne f$ and cannot be obtained from $f$ using transforms
  in the subgroup $\{Id, R\}$.
\end{itemize}
